Carrier phase and amplitude estimation for phase shift keying using pilots and data (Appendix)

### 2012-08-17 paper2.tex
minor edits

\begin{lemma}\label{lem:symmetrisation}

@@

Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$.  Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.  Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.  Then

@@

the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$, it follows that

@@

\begin{lemma} \label{lem:chaining}
For any $\kappa > 0$ there exists $\epsilon > 0$ such that
\[
\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa,
\]
where $X(\lambda)$ is the Gaussian process defined in~\eqref{eq:ZpsiCondGaussProc}.
\end{lemma}
\begin{IEEEproof}
Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
\begin{equation}\label{eq:supZCK1} 
\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}
\end{equation}
where $K_1$ is a finite, positive constant, and
\begin{equation}\label{eq:Cepsdefn}
C_\epsilon = \frac{2}{L}\epsilon\sum_{i\in P \cup D}R_i^2 + \frac{4}{L}\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
\end{equation}
where $I_\epsilon$ is the indicator function
\[
I_{\epsilon}(x) = \begin{cases}
1, & x \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon] \\
0, & \text{otherwise},
\end{cases}
\] 
and $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  Now,
\[
\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
\]
Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
\begin{align*}
\expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
&= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
\end{align*}
where the function
\[
g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
\]
is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.  From the definition of $I_\epsilon$,
\begin{align*}
\expect R_i^2 I_\epsilon(\Phi_i) &= \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \\
&= 2\epsilon \sum_{k=0}^{M-1} (g_2(\psi_k) + o(1))
\end{align*}
where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.  Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$. Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive, it follows that
\[
\expect \sqrt{C_\epsilon} \leq  \sqrt{\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
\]
by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
\[
\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
\]
Choosing 
\[
\epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
\] 
completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
\end{IEEEproof}

@@

\begin{lemma}
Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
\end{lemma}
\begin{IEEEproof}
We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
 \[
 c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) }
 \]
 and 
 \[
 c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
 \]
 As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
 \[
 \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
 \]
 for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let $S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}$. Now $c_2 = c_3 + c_4$ where 
 \[
 c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}
 \]
 and
 \[
 c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
 \]
 If $i \notin S$ then $\round{\lambda + \Phi_i} = \round{\Phi_i}$ for all $\sabs{\lambda} < \epsilon$ and therefore
 \begin{align*}
 \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
 &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
 &\leq \sabs{\lambda} < \epsilon
 \end{align*}
and so, 
\[
c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
\]
Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
\[
c_3 \leq 2 \sum_{i \in D, i \in C} R_i \leq 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i).
\]
Now
\begin{align*}
d_2(\lambda, 0) &= c_1 + c_2 \\
&\leq c_1 + c_3 + c_4 \\
&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i)
\end{align*}
as required
\end{IEEEproof}

>>>>

\begin{lemma}\label{lem:symmetrisation} (Symmetrisation)

@@

Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$.  Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.  Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.  Now,

@@

the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$,

@@

\begin{lemma} \label{lem:chaining}
For any $\kappa > 0$ there exists $\epsilon > 0$ such that
\[
\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa,
\]
where $X(\lambda)$ is the Gaussian process defined in~\eqref{eq:ZpsiCondGaussProc}.
\end{lemma}
\begin{IEEEproof}
Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
\begin{equation}\label{eq:supZCK1} 
\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}
\end{equation}
where $K_1$ is a finite, positive constant, and
\begin{equation}\label{eq:Cepsdefn}
C_\epsilon = \frac{2}{L}\epsilon\sum_{i\in P \cup D}R_i^2 + \frac{4}{L}\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
\end{equation}
where $I_\epsilon$ is the indicator function
\[
I_{\epsilon}(x) = \begin{cases}
1, & x \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon] \\
0, & \text{otherwise},
\end{cases}
\] 
and $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  Now,
\[
\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
\]
Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
\begin{align*}
\expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
&= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
\end{align*}
where the function
\[
g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
\]
is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.  From the definition of $I_\epsilon$,
\begin{align*}
\expect R_i^2 I_\epsilon(\Phi_i) &= \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \\
&= 2\epsilon \sum_{k=0}^{M-1} (g_2(\psi_k) + o(1))
\end{align*}
where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.  Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$. Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive,
\[
\expect \sqrt{C_\epsilon} \leq  \sqrt{\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
\]
by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
\[
\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
\]
Choosing 
\[
\epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
\] 
completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
\end{IEEEproof}

@@

\begin{lemma}\label{lem:epslmlemma}
Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
\end{lemma}
\begin{IEEEproof}
We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
 \[
 c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) }
 \]
 and 
 \[
 c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
 \]
 As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
 \[
 \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
 \]
 for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let $S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}$. Now $c_2 = c_3 + c_4$ where 
 \[
 c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}
 \]
 and
 \[
 c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
 \]
 If $i \notin S$ then $\round{\lambda + \Phi_i} = \round{\Phi_i}$ for all $\sabs{\lambda} < \epsilon$ and therefore
 \begin{align*}
 \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
 &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
 &\leq \sabs{\lambda} < \epsilon
 \end{align*}
and so, 
\[
c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
\]
Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
\[
c_3 \leq 2 \sum_{i \in D, i \in C} R_i \leq 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i).
\]
Now
\begin{align*}
d_2(\lambda, 0) &= c_1 + c_2 \\
&\leq c_1 + c_3 + c_4 \\
&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i)
\end{align*}
as required
\end{IEEEproof}
